---
title: "Chapter 4 - Sampling Distributions"
format: 
  html:
    self-contained: true
    code-fold: true
---

```{r}
#| message: false
#| echo: false

library(resampledata3)
library(tidyverse)
```

## 4.1
Consider the population {1, 2, 5, 6, 10, 12}. Find (and plot) the sampling distribution of medians for samples of size 3 without replacement. Compare the median of the population to the mean of the medians.

```{r}
p <- c(1, 2, 4, 6, 10, 12)

p_sampling_dist <- map_dbl(1:10^4, \(x) sample(p, 3, replace = FALSE) |> median())
```

Plot of the sampling distribution of the medians
```{r}
ggplot() +
  geom_histogram(aes(x=p_sampling_dist), binwidth = 3.4) 
```
```{r}
ggplot() +
  stat_ecdf(aes(x=p_sampling_dist))
```

Mean of the medians
```{r}
p_sampling_dist |> mean()
```

Median of the population
```{r}
median(p)
```

## 4.2
Consider the population {3, 6, 7, 9, 11, 14} For samples of size 3 without replacement, find (and plot) the sampling distribution of the minimum. What is the mean of the sampling distribution? The statistic is an estimate of some parameter – what is the value of that parameter?

```{r}
p4.2 <- c(3, 6, 7, 9, 11, 14)
p4.2_sampling_dist <- map_dbl(1:10^4, \(x) sample(p4.2, 3, replace = FALSE) |> min())
```

Plot of the sampling distribution of the minimum
```{r}
ggplot() +
  geom_density(aes(x=p4.2_sampling_dist))
```
Mean of the sampling distribution
```{r}
mean(p4.2_sampling_dist)
```

The statistic, estimate of the parameter, is 4.8.

## 4.3
Let $A$ denote the population ${1, 3, 4, 5}$ and $B$ the population ${5, 7, 9}$. Let $X$ be a random value from $A$, and $Y$ be a random value from $B$.

a. Find the sampling distribution of $X+Y$.
```{r}
A <- c(1, 3, 4, 5)
B <- c(5, 7, 9)
n <- 3

ab_sampling_dist_nr <- map_dbl(1:10^4, \(x) sample(c(A, B), n, replace = FALSE) |> mean())
ab_sampling_dist_r <- map_dbl(1:10^4, \(x) sample(c(A, B), n, replace = TRUE) |> mean())
```

Plot of the sampling distributions

```{r}
df <- data.frame(r = ab_sampling_dist_r, nr = ab_sampling_dist_nr)
df |> 
  pivot_longer(cols = everything(), names_to = "sampling_method", values_to = "values") |> 
  ggplot(aes(x = values, color=sampling_method)) +
  geom_density()
  
```

b. In this example, does the sampling distribution depend on whether you sample with or without replacement? Why or why not?
For the most frequent number, it appears there's no difference between sampling with replacement compared with sampling with no replacement.

c. Compute the mean of the values for each of $A$ and $B$. Compute the mean of the values in the sampling distribution of $A+B$. How are the means related?

```{r}
list(mean_a=mean(A), mean_b=mean(B), mean_ab_r=mean(ab_sampling_dist_r), mean_ab_nr=mean(ab_sampling_dist_nr))
```
It appears that the mean of $A+B$ is the mean of the mean of the two groups.

d. Suppose you draw a random value from $A$ and a random value from $B$. What is the probability that the sum is 13 or larger?

```{r}
df <- expand.grid(A, B) |> 
  as.data.frame()

df |> 
  rename(a=Var1, b=Var2) |> 
  mutate(s=a+b) |> 
  group_by(s) |> 
  tally() |> 
  mutate(prop=round(n/sum(n), 3))
```
```{r}
0.083*2
```

## 4.4
Consider the population ${3, 5, 6, 6, 8, 11, 13, 15, 19, 20}$.

a. Compute the mean and standard deviation and create a dot plot of its distribution.

```{r}
p4.4 <- c(3, 5, 6, 6, 8, 11, 13, 15, 19, 20) 
list(mean=mean(p4.4), sd=sd(p4.4))
```
```{r}
ggplot() +
  geom_dotplot(aes(x=p4.4))
```



